`r params$doc_title`

### 2022-11-14 scripts/MSc.Rmd
sim no_instruments, sim beta ~N(mu=beta, se=se_n)

---
title: "`r params$doc_title`"
author: | 
        | christoph.reich@med.uni-heidelberg.de 
        | lorenzo@imbi.uni-heidelberg.de
        | 
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
params:
  date: !r Sys.Date()
  doc_title: "MSc Default Title"
output: 
  html_document:
    theme: paper
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 4
    df_print: kable
    latex_engine: xelatex
#bibliography: ../references/project01.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
figurespath <- paste0("../reports/", format(Sys.time(), "%Y%m%d"), '_figures/')
knitr::opts_chunk$set(fig.path = figurespath)  # reference is script! also creates new dirs

library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(scales)
library(ggthemes)
#library(biomaRt)  # different script
library(kableExtra)
library(MendelianRandomization)
library(TwoSampleMR)
library(mr.raps)
library(MRMix)
library(MRPRESSO)
library(penalized)
library(conflicted)
source("MR_lasso.R")  # Slob&Burgess 2020
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

@@

```{r}
harmonised_dat4TwoSampleMR <- harmonised_dat %>% 
  mutate(id.exposure = "BMI", id.outcome = "GBC")  # https://mrcieu.github.io/TwoSampleMR/articles/perform_mr.html

@@

res %>% 
  kable(digits = 3, 
        caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI"
        )

#res.selected_methods <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR, 
#                                        method_list = c("mr_egger_regression", "mr_ivw"))
```

@@

```{r}
mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
est <- tibble(mr_methods=mr_methods, estimate=NA, se.estimate=NA, time=NA)
numIV <- nrow(my_data_harm)

@@

est %>% 
  mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
  kable(digits=3, caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI")
```

>>>>

---
title: "`r params$doc_title`"
author: | 
        | christoph.reich@med.uni-heidelberg.de 
        | lorenzo@imbi.uni-heidelberg.de
        | 
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
params:
  date: !r Sys.Date()
  doc_title: "MSc Default Title"
output: 
  html_document:
    code_folding: hide
    theme: paper
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 4
    df_print: kable
    latex_engine: xelatex
#bibliography: ../references/project01.bib
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
figurespath <- paste0("../reports/", format(Sys.time(), "%Y%m%d"), '_figures/')
knitr::opts_chunk$set(fig.path = figurespath)  # reference is script! also creates new dirs

library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(ggthemes)
library(scales)
library(ggthemes)
#library(biomaRt)  # different script
library(kableExtra)
library(MendelianRandomization)
library(TwoSampleMR)
library(mr.raps)
library(MRMix)
library(MRPRESSO)
library(penalized)
library(conflicted)
source("MR_lasso.R")  # Slob&Burgess 2020
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

@@

```{r TwoSampleMR-analysis}
harmonised_dat4TwoSampleMR <- harmonised_dat %>% 
  mutate(id.exposure = "BMI", id.outcome = "GBC")  # https://mrcieu.github.io/TwoSampleMR/articles/perform_mr.html

@@

res %>% 
  kable(digits = 3, 
        caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI"
        ) %>% 
  kable_classic(full_width = FALSE, html_font = "Cambria")
```

```{r scatter-TwoSampleMR, warning=FALSE, message=FALSE}
# mr_method_list()
res.selected_methods <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR, 
                                        method_list = c("mr_egger_regression", "mr_ivw"))
p1 <- mr_scatter_plot(res.selected_methods, harmonised_dat4TwoSampleMR)
p1[[1]] #+
  #ggthemes::theme_few()

# SAVE
filename.mr.scatter <- paste0("../output/plots/", Sys.Date(), "_figure_mrscatter.png")
ggsave(filename = filename.mr.scatter, plot = p1[[1]], 
         #width = 10, height = 6, 
         #units = "in"  # default
         )
```

```{r forestplot-TwoSampleMR, fig.width=5, fig.height=24, warning=FALSE}
res_single <- TwoSampleMR::mr_singlesnp(harmonised_dat4TwoSampleMR, all_method = c("mr_egger_regression", "mr_ivw"))
p2 <- mr_forest_plot(res_single,  exponentiate = FALSE)
p2[[1]]

#SAVE
filename.mr.forest <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest.png")
ggsave(filename = filename.mr.forest, plot = p2[[1]], 
       width = 5, 
       height = 24, 
         #units = "in"  # default
         )
```

```{r funnel-plot}
p4 <- mr_funnel_plot(res_single)
p4[[1]] +
  ggthemes::theme_few()
```

@@

```{r MendelianRandomization-analysis, warning=FALSE}
mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
est <- tibble(mr_methods=mr_methods, estimate=NA, se.estimate=NA, time=NA)
numIV <- nrow(my_data_harm)

@@

est %>% 
  mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
  kable(digits=3, caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI") %>% 
  kable_classic(full_width = FALSE, html_font = "Cambria")
```

## Simulation

```{r simulatesimulatesimulate}
## try starting with n=50.000 -> selected 9 IVs
set.seed(1234)

## initialize
nsim <- 100
se_update_BMI <- c()
se_update_GBC <- c()

est <- tibble(no_sim = 1:nsim)
#est <- matrix(NA, 
#              nrow = nsim, 
#              ncol = 2*length(n)  # theta_sim and theta_se_sim
#              ) 
#colnames(est) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))

## START SIM
T0 = proc.time()[3]
for (i in 1:length(colnames_pval)) {
  T0 = proc.time()[3]
  threshold <- 5*10^-6
  numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
  if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
    next
  }

  # FILTER DATA
  sim_dat <- my_data_harm %>% 
    filter(!!sym(colnames_pval[i]) < threshold) %>% 
    select(SNP, eaf.bmi, beta.bmi, p.value.bmi, eaf.gbc, beta.gbc, p.value.gbc)

  # get simulated se_{sample size}
  for (j in 1:nrow(sim_dat)) {
    # as done before in section "simulating on empirical data"
    se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
    se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
  }

  sim_dat <- sim_dat %>%
      bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))

  # simulate beta values from normal_distribution
  for (k in 1:nsim) {
    sim_dat <- sim_dat %>% 
      mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
      mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y)))

    # PERFORM MR_analysis
    mr.obj = MendelianRandomization::mr_input(
          bx = sim_dat$beta.bmi_sim_norm, 
          bxse = sim_dat$se_update_BMI, 
          # outcome
          by = sim_dat$beta.gbc_sim_norm, 
          byse = sim_dat$se_update_GBC, 
          snps = sim_dat$SNP,
          exposure = "Body mass index",
          outcome = "Gallbladder Cancer"
        ) 
    # CONMIX (as reference)
    res = MendelianRandomization::mr_conmix(mr.obj)
    colname_theta <- paste0("theta_sim_", n[i])
    colname_theta_se <- paste0("theta_se_sim_", n[i])
    ## store results
    est[k, colname_theta] <- res$Estimate
    CIlength = res$CIUpper-res$CILower
    if (length(CIlength)>1) print("conmix multimodal")
    est[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
    rm(res)
  }
  print(paste0("|||-----------------------Run finished for sample size: ", n[i], " -----------------------|||"))
}
T1 = proc.time()[3]
timediff = T1-T0
```
